---
title: "sim_medals"
output: html_document
date: "2023-09-24"
---
```{r}
library(tidyr)
```


```{r}
# function simulate_medal(team_country_qualified_individuals, means_df, stddevs_df, 36_athletes, gender)
top12teams <- read.csv("data/team_country_qualified_individuals.csv")
means_df <- read.csv("data/means_per_app.csv") #means_df
stddevs_df <- read.csv("data/stddevs_per_app.csv") #stddevs_df
qual36 <- read.csv("data/mens_36_athletes.csv")
gender <- "m"

means_df <- means_df[means_df$Gender == gender, ]
stddevs_df <- stddevs_df[stddevs_df$Gender == gender, ]
top12teams <- top12teams[top12teams$Gender == gender, ]


if(gender=="m"){
  apps <- c("VT", "FX", "HB", "PB", "PH", "SR")
}else{
  apps <- c("VT", "BB", "UB", "FX")
}

```



# Create QUalifying Samples
```{r}
get_samples_df <- function(meansdf, stddevsdf){
  long_means <- pivot_longer(meansdf,
                             cols=c("VT", "BB", "UB", "FX", "HB", "PB", "PH", "SR"),
                             names_to="App", values_to="mean")
  long_stddevs <- pivot_longer(stddevsdf,
                               cols=c("VT", "BB", "UB", "FX", "HB", "PB", "PH", "SR"),
                               names_to="App", values_to="stddev")
  long_data <- merge(x=long_means, y=long_stddevs, by=c("ID", "Gender", "Country", "App"), all=TRUE)
  
  samples_df <- long_data[, c("ID", "Gender", "Country", "App")]
  samples_df$Score <- rnorm(length(long_data$ID), mean=long_data$mean, sd=long_data$stddev)
  samples_df <- pivot_wider(samples_df, names_from="App", values_from="Score")

  return(samples_df)
}
samples_df <- get_samples_df(means_df, stddevs_df)
```

# Team Qualifications
```{r}
means_q <- means_df %>%
  filter(ID %in% top12teams$ID)

means_q$Total_Mean <- rowSums(means_q[, c("VT", "BB", "UB", "FX", "HB", "PB", "PH", "SR")], na.rm = TRUE)

means_q <- means_q %>%
  mutate(across(VT:SR, ~ifelse(is.na(.), 0, .)))

# Selecting top 4 means to compete in qualifying
means_q2 <- means_q %>%
  pivot_longer(cols = VT:SR, names_to = "Apparatus", values_to = "Mean_Score")

top_4_athletes_per_apparatus <- means_q2 %>%
  group_by(Country, Gender, Apparatus) %>%
  arrange(desc(Mean_Score)) %>%
  slice_head(n = 4) %>%
  ungroup() %>%
  filter(!(Gender == "m" & (Apparatus %in% c("BB", "UB")) 
           & Mean_Score == 0.000000)) %>%
  filter(!(Gender == "w" & (Apparatus %in% c("HB", "PB", "PH", "SR")) 
           & Mean_Score == 0.000000))

samples_longer <- samples_df %>%
  pivot_longer(cols = VT:SR, names_to = "Apparatus", values_to = "Final_Score")

# Left join top_4_athletes_per_apparatus with samples_longer by ID and Apparatus
top_4_athletes_per_apparatus <- top_4_athletes_per_apparatus %>%
  left_join(samples_longer, by = c("ID", "Apparatus")) %>%
  mutate(Final_Score = ifelse(is.na(Final_Score), 0, Final_Score)) %>%
  rename(Gender = Gender.x, Country = Country.x) %>%
  select(-Mean_Score, -Country.y, -Gender.y)  # Remove the Mean_Score column if you no longer need it

# Scores for top countries
top_df <- top_4_athletes_per_apparatus %>%
  filter(Gender == gender) %>%
  group_by(Country, Apparatus) %>%
  arrange(desc(Final_Score))

# Slice the top 3 rows for each group (dropping the 4th highest Final_Score)
filtered_df <- top_df %>%
  slice(1:3) %>%
  group_by(Country) %>%
  summarize(Final_Score = sum(Final_Score, na.rm = TRUE)) %>%
  arrange(desc(Final_Score)) %>%
  mutate(Gender = "m")

# Select the top 8 countries with the highest Final_Score
top_8_countries <- head(filtered_df, n = 8)
qual_teams <- top_8_countries$Country 
```


# AA QUalifications
```{r}

```


# Event Qualifications
```{r}

```


# SIMULATE FINALS PREP

```{r}
# the output df, tallying the number of each medal for each country
country_medals <- data.frame(
  Country = unique(means_df$Country),
  Firsts = 0, Seconds = 0, Thirds = 0, Fourths = 0,
  Fifths = 0, Sixths = 0, Sevenths = 0, Eighths = 0
)

# use as dictionary to get the country of each ID (player)
ID_country_dict <- means_df[, c("ID", "Country")]
ID_country_dict <- ID_country_dict[!duplicated(ID_country_dict$ID, fromLast=T), ]
```


# Simulate Team Finals (M)
qual_teams <- read.csv("data/finaled_teams.csv")
```{r}
country_scores <- data.frame(
  country = qual_teams,
  scores = rep(0, 8), 
  nas = rep(0, 8)
)
for(country in qual_teams){
  n_nas <- 0
  athlete_names <- top12teams$ID[top12teams$Country==country]
  
  for(app in apps){
    curr_df <- means_df[means_df$ID %in% athlete_names,]
    top3 <- order(-curr_df[, app])[1:3]
    top3_id <- curr_df$ID[top3]
    for(id in top3_id){
      score <- rnorm(1, 
                     mean=means_df[means_df$ID==id, app],
                     sd=stddevs_df[stddevs_df$ID==id, app]
                  )
      if(!is.na(score)){ # only add if not na, and tally up nas
        country_scores$scores[country_scores$country == country] <- country_scores$scores[country_scores$country == country] + score
      }
      else{
        n_nas <- n_nas + 1
      }
    }
  }
  country_scores$nas[country_scores$country == country] <- n_nas
}

#assigning medals
top8_countries <- country_scores$country[order(-country_scores$scores)[1:8]]
country_medals$Firsts[country_medals$Country == top8_countries[1]] = country_medals$Firsts[country_medals$Country == top8_countries[1]] + 1
country_medals$Seconds[country_medals$Country == top8_countries[2]] = country_medals$Seconds[country_medals$Country == top8_countries[2]] + 1
country_medals$Thirds[country_medals$Country == top8_countries[3]] = country_medals$Thirds[country_medals$Country == top8_countries[3]] + 1
country_medals$Fourths[country_medals$Country == top8_countries[4]] = country_medals$Fourths[country_medals$Country == top8_countries[4]] + 1
country_medals$Fifths[country_medals$Country == top8_countries[5]] = country_medals$Fifths[country_medals$Country == top8_countries[5]] + 1
country_medals$Sixths[country_medals$Country == top8_countries[6]] = country_medals$Sixths[country_medals$Country == top8_countries[6]] + 1
country_medals$Sevenths[country_medals$Country == top8_countries[7]] = country_medals$Sevenths[country_medals$Country == top8_countries[7]] + 1
country_medals$Eighths[country_medals$Country == top8_countries[8]] = country_medals$Eighths[country_medals$Country == top8_countries[8]] + 1
```

```{r}
print(country_scores)
```

```{r}
curr_df
```


# Simulate AA Finals(M)
qual_aa <- read.csv("data/finaled_aas.csv")
```{r}
aa_scores <- data.frame(
  ID = as.character(qual_aa[1, 2:25]),
  scores = rep(0, 24),
  nas = rep(0, 24)
)

for(i in 2:25){
  n_nas <- 0
  id <- qual_aa[1, i]
  for(app in apps){
      score <- rnorm(1,
                     mean=means_df[means_df$ID==id, app],
                     sd=stddevs_df[stddevs_df$ID==id, app]
                  )
      if(!is.na(score)){
        aa_scores$scores[aa_scores$ID == id] <- aa_scores$scores[aa_scores$ID == id] + score
      }
      else{
        n_nas <- n_nas + 1
      }
  }
  aa_scores$nas[aa_scores$ID == id] <- n_nas
}

#assigning medals
top8_i <- aa_scores$ID[order(-aa_scores$scores)[1:8]]
top8_countries <- as.character(sapply(top8_i, function(i) ID_country_dict$Country[ID_country_dict$ID == i]))
country_medals$Firsts[country_medals$Country == top8_countries[1]] = country_medals$Firsts[country_medals$Country == top8_countries[1]] + 1
country_medals$Seconds[country_medals$Country == top8_countries[2]] = country_medals$Seconds[country_medals$Country == top8_countries[2]] + 1
country_medals$Thirds[country_medals$Country == top8_countries[3]] = country_medals$Thirds[country_medals$Country == top8_countries[3]] + 1
country_medals$Fourths[country_medals$Country == top8_countries[4]] = country_medals$Fourths[country_medals$Country == top8_countries[4]] + 1
country_medals$Fifths[country_medals$Country == top8_countries[5]] = country_medals$Fifths[country_medals$Country == top8_countries[5]] + 1
country_medals$Sixths[country_medals$Country == top8_countries[6]] = country_medals$Sixths[country_medals$Country == top8_countries[6]] + 1
country_medals$Sevenths[country_medals$Country == top8_countries[7]] = country_medals$Sevenths[country_medals$Country == top8_countries[7]] + 1
country_medals$Eighths[country_medals$Country == top8_countries[8]] = country_medals$Eighths[country_medals$Country == top8_countries[8]] + 1
```


# Simulate Event Finals
qual_ind <- read.csv("data/finaled_events.csv")
```{r}
get_event_final_scores <- function(gender, app, IDs){
  out_df <- data.frame(
    ID = IDs,
    gender = rep(gender, length(IDs)),
    app = rep(app, length(IDs)),
    score = rep(0, length(IDs))
  )
  out_df$Score <- sapply(out_df$ID, 
                         function(id) rnorm(1,
                                            mean=means_df[means_df$ID==id, app],
                                            sd=stddevs_df[stddevs_df$ID==id, app]
                                            )
                         )
  return(out_df)
}


for(i in 1:dim(qual_ind)[1]){
  curr_event <- qual_ind[i, ] 
  scores <- get_event_final_scores(curr_event$Gender, curr_event$App, as.character(curr_event[, 3:ncol(curr_event)]))
  
  top8_i <- scores$ID[order(-scores$score)[1:8]]
  top8_countries <- as.character(sapply(top8_i, function(i) ID_country_dict$Country[ID_country_dict$ID == i]))
  country_medals$Firsts[country_medals$Country == top8_countries[1]] = country_medals$Firsts[country_medals$Country == top8_countries[1]] + 1
  country_medals$Seconds[country_medals$Country == top8_countries[2]] = country_medals$Seconds[country_medals$Country == top8_countries[2]] + 1
  country_medals$Thirds[country_medals$Country == top8_countries[3]] = country_medals$Thirds[country_medals$Country == top8_countries[3]] + 1
  country_medals$Fourths[country_medals$Country == top8_countries[4]] = country_medals$Fourths[country_medals$Country == top8_countries[4]] + 1
  country_medals$Fifths[country_medals$Country == top8_countries[5]] = country_medals$Fifths[country_medals$Country == top8_countries[5]] + 1
  country_medals$Sixths[country_medals$Country == top8_countries[6]] = country_medals$Sixths[country_medals$Country == top8_countries[6]] + 1
  country_medals$Sevenths[country_medals$Country == top8_countries[7]] = country_medals$Sevenths[country_medals$Country == top8_countries[7]] + 1
  country_medals$Eighths[country_medals$Country == top8_countries[8]] = country_medals$Eighths[country_medals$Country == top8_countries[8]] + 1
}
```

```{r}
write.csv(country_medals, "data/medalling_outputs.csv", row.names=F)
```

```{r}
colSums(country_medals[, 2:9])
```



To-Do: load qual_team and qual_ind in properly, and output the medals df






